Plant trait cleaning and combining

### 2024-05-22 code/L1/L1_LastStep_DiversityInputObjects.Rmd
project PAM and update object names

## Turn PAM into matrix
```{r}
plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
```

## Add rownames
```{r}
# Get the number of rows in the matrix
num_rows_plant <- nrow(plants_PAM_matrix)

@@

# Assign row names to the matrix
rownames(plants_PAM_matrix) <- row_names_plant

column_names_plant <- colnames(plants_PAM_matrix)
```

```{r}
# Get the number of rows in the matrix
num_rows_frugivore <- nrow(frugivore_PAM_matrix)

@@

# Assign row names to the matrix
rownames(frugivore_PAM_matrix) <- row_names_frugivore

column_names_frugivore <- colnames(frugivore_PAM_matrix)
```

@@

## Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final <- plants_PAM_matrix[, colSums(plants_PAM_matrix) != 0]
PAM_frugivore_site_final <- frugivore_PAM_matrix[, colSums(frugivore_PAM_matrix) != 0]

>>>>

## Turn PAM into matrix
```{r}
plant_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
```

## Project PAM
```{r}
# check dimensions
dim(plant_PAM_matrix)
dim(frugivore_PAM_matrix)
```

```{r}
# convert PAM matrix to dataframe
plant_PAM_df <- as.data.frame(plant_PAM_matrix)
frugivore_PAM_df <- as.data.frame(frugivore_PAM_matrix)
```

```{r}
# check dimensions
dim(plant_PAM_df)
dim(frugivore_PAM_df)
```

```{r}
# Convert to sf object and assign crs
plant_PAM.sf <- st_as_sf(plant_PAM_df, coords = c("Longitude(x)", "Latitude(y)"), crs = 4326)
frugivore_PAM.sf <- st_as_sf(frugivore_PAM_df, coords = c("Longitude(x)", "Latitude(y)"), crs = 4326)
```

```{r}
# check dimensions
dim(plant_PAM.sf)
dim(frugivore_PAM.sf)
```

```{r}
# project to crs to match taxonomic diversity
plant_PAM.sf <- st_transform(plant_PAM.sf, 5389)
frugivore_PAM.sf <- st_transform(frugivore_PAM.sf, 5389)
```

```{r}
# check units
st_crs(plant_PAM.sf)
st_crs(frugivore_PAM.sf)
```

```{r}
# convert sf to dataframe and retain lat & long as first two columns
## extract coordinates
plant_coords <- st_coordinates(plant_PAM.sf)
frugivore_coords <- st_coordinates(frugivore_PAM.sf)

## convert sf to dataframe
plant_PAM.sf_df <- as.data.frame(st_drop_geometry(plant_PAM.sf))
frugivore_PAM.sf_df <- as.data.frame(st_drop_geometry(frugivore_PAM.sf))

# add coordinates as separate columns
plant_PAM.sf_df$Longitude <- plant_coords[, 1]
plant_PAM.sf_df$Latitude <- plant_coords[, 2]

frugivore_PAM.sf_df$Longitude <- frugivore_coords[, 1]
frugivore_PAM.sf_df$Latitude <- frugivore_coords[, 2]

# rearrange columns to make Longitude and Latitude the first two columns
plant_PAM.sf_df <- plant_PAM.sf_df[, c("Longitude", "Latitude", setdiff(names(plant_PAM.sf_df), c("Longitude", "Latitude")))]
frugivore_PAM.sf_df <- frugivore_PAM.sf_df[, c("Longitude", "Latitude", setdiff(names(frugivore_PAM.sf_df), c("Longitude", "Latitude")))]
```

```{r}
#check str of dataframes
plant_PAM.sf_df[1:4, 1:4]
frugivore_PAM.sf_df[1:4, 1:4]
```

```{r}
# convert back to matrix and make sure all columns are numeric
plant_PAM.sf_matrix <- as.matrix(plant_PAM.sf_df)
frugivore_PAM.sf_matrix <- as.matrix(frugivore_PAM.sf_df)

plant_PAM.sf_matrix <- apply(plant_PAM.sf_matrix, 2, as.numeric)
frugivore_PAM.sf_matrix <- apply(frugivore_PAM.sf_matrix, 2, as.numeric)
```

```{r}
# check dimensions
dim(plant_PAM.sf_matrix)
dim(frugivore_PAM.sf_matrix)
```

## Add rownames
```{r}
# Get the number of rows in the matrix
num_rows_plant <- nrow(plant_PAM.sf_matrix)

@@

# Assign row names to the matrix
rownames(plant_PAM.sf_matrix) <- row_names_plant

column_names_plant <- colnames(plant_PAM.sf_matrix)
```

```{r}
# Get the number of rows in the matrix
num_rows_frugivore <- nrow(frugivore_PAM.sf_matrix)

@@

# Assign row names to the matrix
rownames(frugivore_PAM.sf_matrix) <- row_names_frugivore

column_names_frugivore <- colnames(frugivore_PAM.sf_matrix)
```

@@

## Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final <- plant_PAM.sf_matrix[, colSums(plant_PAM.sf_matrix) != 0]
PAM_frugivore_site_final <- frugivore_PAM.sf_matrix[, colSums(frugivore_PAM.sf_matrix) != 0]
